---
title: "Bond Disbursement"
author: "Allen Church"
date: "October 29, 2019"
output: html_document
---
- Look at PPG bond disbursement (DT.DIS.PBND.CD) for LAC, as it's in top 4 aggregate.
- Look at Foreign direct investment, net inflows in reporting economy (BX.KLT.DINV.CD.DT) and see how it relates to PPG bond DIS
- Look at Exports of goods, services and primary income (BX.GSR.TOTL.CD) and see how it relates to PPG bond DIS

Fetch PPG bond disbursement 
```{r}
library(wbstats)

#Search indicators on Databank
wbsearch(pattern = "PPG, bond")

#Create vector for Latin America and Carribean
LAC <- c("LAC")

#Specify PPG bond disbursement
ppg_bond_dis <- ("DT.DIS.PBND.CD")

#Fetch data from WB API, specifying country, indicator, start/end dates and put into dataframe
bond_data_LAC <- wb(country = LAC, indicator = ppg_bond_dis, startdate = 2014, enddate = 2018, POSIXct = TRUE)
df <- data.frame(bond_data_LAC)
View(df)
```


Clean and format data
```{r}
#Subset ppg bond disbursement into dataframe, format units to thousands, subset
ppg_bond_dis <- subset(df, indicatorID == "DT.DIS.PBND.CD")

#Change units to billions USD
ppg_bond_dis$disbursement <- ppg_bond_dis$value/1000000000
ppg_bond_dis <- subset(ppg_bond_dis, select = c("date", "indicator", "country", "disbursement"))
View(ppg_bond_dis)
```


Add exports of goods, services, and primary income and create new dataframe
```{r}
#Create vector to include PPG Bond disbursement and export of GSI
ppg_dis_export <- c("DT.DIS.PBND.CD", "BX.GSR.TOTL.CD")

#Add new indicator to bond_export_df and view
bond_export_LAC <- wb(country = LAC, indicator = ppg_dis_export, startdate = 2014, enddate = 2018, POSIXct = TRUE)
bond_export_df <- data.frame(bond_export_LAC)

#Change units to billions USD
bond_export_df$Values <- bond_export_df$value/1000000000
bond_export_df <- subset(bond_export_df, select = c("date", "indicator", "country", "Values"))
View(bond_export_df)
```

Graph PPG bond disbursement with export GSI
```{r}
library(ggplot2)

ggplot(data = bond_export_df, mapping = aes(x=date, y=Values, color=indicator)) +
  geom_line()

```

List of LAC countries
```{r}
LAC_list <- c("ARG", "BRA", "COL", "CRI", "DOM", "ECU", "GTM", "HND", "MEX", "PRY")

#Fetch data from WB API, specifying country, indicator, start/end dates and put into dataframe
bond_data_LAC_list <- wb(country = LAC_list, indicator = "DT.DIS.PBND.CD", startdate = 2014, enddate = 2018, POSIXct = TRUE)
df_LAC_list <- data.frame(bond_data_LAC_list)
df_LAC_list$Values <- df_LAC_list$value/1000000000
df_LAC_list <- subset(df_LAC_list, select = c("date", "indicator", "country", "Values"))
View(df_LAC_list)
```

Graph attempt 2
```{r}
ggplot(data = df_LAC_list, mapping = aes(x=date, y=Values, group=country, color=country)) +
  geom_point() + 
  geom_line() 

```

