---
output: pdf_document
---

Allen Church
Chapter 2 - Exercise 2
Accelerated Statistics for Public Policy
9/16/19

Load data and rename to data1. View data to look at dataset.
```{r}
load("/Users/allenchurch/Ch2_Exercise2_Olympics.RData")
data1 <- dta
```

2a. Summarize medals, athletes, and GDP columns.

```{r}
summary(data1$medals)
```

```{r}
summary(data1$athletes)
```
```{r}
summary(data1$GDP)
```

2b. List first 5 observations for country, year, medals, athletes, and GDP data.

```{r}
##Create cols with 5 column names, subset data1 with cols, and print first 5 rows
cols <- c("country", "year", "medals", "athletes", "GDP")
subset <- data1[cols]
subset[1:5,]
```

2c. How many observations are there for each year?
```{r}
#Create table with year column in data1
table(data1$year)
```

2d. Produce scatterplot of medals and number of athletes. Describe relationship depicted.

The relationship depicted below shows high clustering in the number of athletes that received 0 medals. It also shows that a majority - or close to a majority - of athletes received under 10 medals. Moreover, the number of athletes that received 30 or more medals is incredibly low with a count of 5 athletes. It also shows a positive correlation between number of athletes and number of medals won.

```{r}
plot(data1$medals, data1$athletes, main="Scatterplot of Medals and Number of Athletes",
     xlab="Medals", ylab = "Number of Athletes")
```

2e. Explain any suspicion you might have that other factors could explain the observed relationship between number of athletes and medals.

It seems there is a direct relationship between the number of athletes and the number of medals won by a country. This may cause the independent variable to be endogenous if changes in it are related to factors of the error term.

2f. Create scatterplot of medals and GDP. Briefly describe any clear patterns.

There does not seem to be a clear relationship between GDP and number of medals earned, meaning that GDP is not strongly correlated with number of medals. It also looks as the majority of countries won 15 or less medals, and there is also a high concentration of countries with 0 medals won.

```{r}
plot(data1$medals, data1$GDP, main="Scatterplot of Medals and GDP",
     xlab="Medals", ylab = "GDP")
```

2g. Create scatterplot of medals and population. Briefly describe any clear patterns.

The relationship below shows that population size is not strongly correlated with number of medals won. As seen in the scatterplot below, countries with very large populations won the same, or less, medals than countries with smaller populations.

```{r}
plot(data1$medals, data1$population, main="Scatterplot of Medals and Population",
     xlab="Medals", ylab = "Population")
```

2h. Create scatterplot of medals and temperature. Describe any clear patterns

The relationship below shows that countries with high medal counts - more than 20 - have an average temperature of approximately 40 degrees. Countries with extremely high temperatures - above 60 degrees F - mostly received 0 medals in the games. Since it was the Winter Olympics, this relationship is logical.

```{r}
plot(data1$medals, data1$temp, main="Scatterplot of Medals and Population",
     xlab="Medals", ylab = "Temperature in F")
```

