---
title: "Chapter 3 Exercise 4"
author: "Allen Church"
date: "September 24, 2019"
output:
  word_document: default
  html_document: default
---
4a. For each data set (women and men), create scatterplot of hours worked on y axis and divorce rates on x axis

```{r}
#Load datasets and view
load("C:/Users/WB537822/Desktop/stats1/Ch3_Exercise4_Divorce_rates_Women.RData")
load("C:/Users/WB537822/Desktop/stats1/Ch3_Exercise4_Divorce_rates_Men.RData")

View(Mdata)
View(Wdata)
```

Scatterplot for women dataset
```{r}
plot(Wdata$divorcerate, Wdata$hours, main="Scatterplot of Female Divorce Rate vs. Hours Worked",
     xlab="Divorce Rate", ylab = "Hours Worked", pch=9)
```

Scatterplot for men dataset

```{r}
plot(Mdata$divorcerate, Mdata$hours, main="Scatterplot of Male Divorce Rate vs. Hours Worked",
     xlab="Divorce Rate", ylab = "Hours Worked", pch=9)
```

4b. For each data set estimate OLS regression in which hours worked is regressed on divorce rates. Report estimated regression equation and interpret coefficients. Explain differences in coefficients.

Women dataset OLS regression: The formula for this regression is the dependent variable = hours worked and the independent variable = divorce rate.
```{r}
#Create OLS regression for women dataset
ols_women <- lm(Wdata$hours ~ Wdata$divorcerate)
summary(ols_women)
```

Men dataset OLS regression: The formula for this regression is the dependent variable = hours worked and the independent variable = divorce rate.
```{r}
#Create OLS regression for men dataset
ols_men <- lm(Mdata$hours ~ Mdata$divorcerate)
summary(ols_men)
```

The results between the two datasets above suggest that, since the beta hat 1 for women is 48.28 vs. beta hat 1 for men is 1.798, hours worked and divorce rate have a much higher positive correlation for women.

4c. What are fitted value and residual for men in Germany?
```{r}
#Create a new dataframe only containing men data from Germany
german_men <- subset(Mdata, country == "Germany")
```

Fitted value for men in Germany
```{r}
ols_german_men <- lm(german_men$hours ~ german_men$divorcerate)
ols_german_men$fitted.values
```

Residual for men in Germany
```{r}
summary(ols_german_men)
```
The results above show there are 0 residuals for men in Germany.

4d. Fitted value and residual for women in Spain
```{r}
#Create a new dataframe only containing women data from Spain
spanish_women <- subset(Wdata, country == "Spain")
ols_spanish_women <- lm(spanish_women$hours ~ spanish_women$divorcerate)
```

Fitted value for women in Spain
```{r}
ols_spanish_women$fitted.values
```

Residual for women in Spain
```{r}
summary(ols_spanish_women)
```
The results above show there are 0 residuals for women in Spain.