---
title: "Chapter 4 Exercise 4"
author: "Allen Church"
date: "10/9/2019"
output:
  word_document: default
  html_document: default
---
4. Run simulation code in the initial part of the education and salary question from the exercises in Ch 3 (p 87)
```{r}
## Set model and simulation parameters
Obs       = 100		# Number of observations in each simulation
Reps      = 50		# Number of times we run the simulation
TrueBeta0	= 12000	# "True" beta0 for the simulated
TrueBeta1	= 1000	# "True" beta1 for the simulated
SD 		    = 10000	# The standard deviation of the error. The bigger this is, the larger the average value of epsilon
Ed = 16 * runif(Obs)# Simulate years of education as being between 0 and 16
  # "runif" is a uniform random variable between 0 and 1, with all values having equal probability
CoefMatrix	= matrix(NA, Reps, 2)
SEMatrix    = matrix(NA, Reps, 2)
# Matrix to store our results - for both coefficients and std errors
  # 1st argument is NA, meaning we store "not available" as initial values in the matrix
  # 2nd argument is Reps, meaning the number of rows is equal to number of times we run the simulations
  # 3rd argument is 2 meaning we have 2 columns, one for storing the beta0 estimate and one for storing the beta1 estimate

# Loop: repeat the commands between the brackets multiple times
for (ii in 1:Reps) { 
  Salary 	= TrueBeta0+ TrueBeta1* Ed + SD*rnorm(Obs) 	
  # Generate Salary = beta0 + beta1*Ed + epsilon
  # beta0 is the constant
  # beta1 is the number multiplied by the X variable
  # Epsilon has 2 parts: SD is the standard deviation; the bigger it is, the more epsilon varies.
  # "runif" is a uniform random variable between 0 and 1, with all values having equal probability
  OLS.result = lm(Salary ~ Ed) # Run a regression using simulated values of Y
  CoefMatrix[ii, ]	= coefficients(OLS.result)	 # Put OLS.result coefficients in row ii of CoefMatrix
  SEMatrix[ii, ]  = coef(summary(OLS.result))[, "Std. Error"]
  ## For fun: plot results for each survey
  ## plot(Ed, Salary, pch = 19, col= "darkgreen")		
  ## abline(OLS.result, lwd = 3, col= "darkgreen")
  ## Sys.sleep(0.075)		## Include to slow down calculations so we can see each plot (briefly); not necessary
}							 # This closes the "loop"

c(mean(CoefMatrix[,1]), min(CoefMatrix[,1]), max(CoefMatrix[,1]))  
# Average, min and max of beta_0 estimates

c(mean(CoefMatrix[,2]), min(CoefMatrix[,2]), max(CoefMatrix[,2]))  
# Average, min and max of beta_1 estimates

#
# For use in Chapter 3, #2 part (g):
#
# Kernel Density Plot
plot(density(CoefMatrix[,2]), main = 'Kernel Density Estimate')

summary(OLS.result)
```

t stat = beta 1 hat / standard error of beta 1 hat
tmatrix <- CoefMatrix/SEmatrix
```{r}
#Create a T matrix, which is results from the coefficient matrix divided by results from the standard error matrix
T.matrix = CoefMatrix/SEMatrix

## Minimum t-stats for both coefficients (b0, b1)
c(min(T.matrix[, 1]),  min(T.matrix[, 2]) )

#Print minimum beta 0 t-stat to user
print(paste0("The minimum t-stat for beta 0 is ", min(T.matrix[, 1])))

#Print minimum beta 1 t-stat to user
print(paste0("The minimum t-stat for beta 1 is ", min(T.matrix[, 2])))

## Maximum t-stats for both coefficients (b0, b1)
c(max(T.matrix[, 1]),  max(T.matrix[, 2]) )

#Print maximum beta 0 t-stat to user
print(paste0("The maximum t-stat for beta 0 is ", max(T.matrix[, 1])))

#Print maximum beta 1 t-stat to user
print(paste0("The maximum t-stat for beta 1 is ", max(T.matrix[, 2])))
```
4b. Generate two-sided p values for the coefficient on education for each simulation. What are minimal and maximal values of these p values?
```{r}
#Create two sided p values, and assign degrees of freedom from lm summary in 4a.
deg_freedom <- 98
pvals <- 2*(1-pt(abs(T.matrix),deg_freedom))

#Create P matrix with necessary shape to store results below
P.Matrix = matrix(NA, Reps, 2)

# For loop to assign p vals into the P Matrix
for (ii in 1:Reps) { 
  #OLS.result = lm(Salary ~ Ed) # Run a regression using simulated values of Y
  #CoefMatrix[ii, ]	= coefficients(OLS.result)	 # Put OLS.result coefficients in row ii of CoefMatrix
  #SEMatrix[ii, ]  = coef(summary(OLS.result))[, "Std. Error"]
  P.Matrix = pvals
  
}

## Minimum t-stats for both coefficients (b0, b1)
c(min(P.Matrix[, 1]),  min(P.Matrix[, 2]) )

#Print minimum beta 0 p value to user
print(paste0("The minimum p value for beta 0 is ", min(P.Matrix[, 1])))

#Print minimum beta 1 p value to user
print(paste0("The minimum p value for beta 1 is ", min(P.Matrix[, 2])))

## Maximum p values for both coefficients (b0, b1)
c(max(P.Matrix[, 1]),  max(P.Matrix[, 2]) )

#Print maximum beta 0 p value to user
print(paste0("The maximum p value for beta 0 is ", max(P.Matrix[, 1])))

#Print maximum beta 1 p value to user
print(paste0("The maximum p value for beta 1 is ", max(P.Matrix[, 2])))
```

4c. In what percent of simulations do we reject null hypothesis that Beta education = 0 at alpha = 0.05 level with a two-sided alternative hypothesis?

The maximum p value for Beta 1 above is 0.032. This indicates a 0.032 probability of observing the Beta education coefficient as large as observed if the Beta education coefficient equals 0.

4d. Re-run simulations and set true value of Beta education to 0. Do this for 500 simulations and report what percent of time we reject the null at alpha = 0.05 level with a two-sided alternative hypothesis

```{r}
## Set model and simulation parameters
Obs_1       = 100		# Number of observations in each simulation
Reps_1      = 500		# Number of times we run the simulation
TrueBeta0_1	= 48	# "True" beta0 for the simulated
TrueBeta1_1	= 0	# "True" beta1 for the simulated
SD 		    = 10000	# The standard deviation of the error. The bigger this is, the larger the average value of epsilon
Ed_1 = 16 * runif(Obs)# Simulate years of education as being between 0 and 16
  # "runif" is a uniform random variable between 0 and 1, with all values having equal probability
CoefMatrix_1	= matrix(NA, Reps_1, 2)
SEMatrix_1    = matrix(NA, Reps_1, 2)

# Matrix to store our results - for both coefficients and std errors
  # 1st argument is NA, meaning we store "not available" as initial values in the matrix
  # 2nd argument is Reps, meaning the number of rows is equal to number of times we run the simulations
  # 3rd argument is 2 meaning we have 2 columns, one for storing the beta0 estimate and one for storing the beta1 estimate

# Loop: repeat the commands between the brackets multiple times
for (ii in 1:Reps_1) { 
  Salary_1 	= TrueBeta0_1+ TrueBeta1_1* Ed_1 + SD*rnorm(Obs_1) 	
  # Generate Salary = beta0 + beta1*Ed + epsilon
  # beta0 is the constant
  # beta1 is the number multiplied by the X variable
  # Epsilon has 2 parts: SD is the standard deviation; the bigger it is, the more epsilon varies.
  # "runif" is a uniform random variable between 0 and 1, with all values having equal probability
  OLS.result_1 = lm(Salary_1 ~ Ed_1) # Run a regression using simulated values of Y
  CoefMatrix_1[ii, ]	= coefficients(OLS.result_1)	 # Put OLS.result coefficients in row ii of CoefMatrix
  SEMatrix_1[ii, ]  = coef(summary(OLS.result_1))[, "Std. Error"]
}

T.matrix_1 = CoefMatrix_1/SEMatrix_1

#Create two sided p values, and assign degrees of freedom from lm summary in 4a.
deg_freedom_1 <- 98
pvals_1 <- 2*(1-pt(abs(T.matrix_1),deg_freedom_1))

#Create P matrix with necessary shape to store results below
P.Matrix_1 = matrix(NA, Reps_1, 2)

# For loop to assign p vals into the P Matrix
for (ii in 1:Reps_1) { 
  #OLS.result = lm(Salary ~ Ed) # Run a regression using simulated values of Y
  #CoefMatrix[ii, ]	= coefficients(OLS.result)	 # Put OLS.result coefficients in row ii of CoefMatrix
  #SEMatrix[ii, ]  = coef(summary(OLS.result))[, "Std. Error"]
  P.Matrix_1 = pvals_1
  
}

## Minimum t-stats for both coefficients (b0, b1)
c(min(P.Matrix_1[, 1]),  min(P.Matrix_1[, 2]) )

#Print minimum beta 0 p value to user
print(paste0("The minimum p value for beta 0 is ", min(P.Matrix_1[, 1])))

#Print minimum beta 1 p value to user
print(paste0("The minimum p value for beta 1 is ", min(P.Matrix_1[, 2])))

## Maximum p values for both coefficients (b0, b1)
c(max(P.Matrix_1[, 1]),  max(P.Matrix_1[, 2]) )

#Print maximum beta 0 p value to user
print(paste0("The maximum p value for beta 0 is ", max(P.Matrix_1[, 1])))

#Print maximum beta 1 p value to user
print(paste0("The maximum p value for beta 1 is ", max(P.Matrix_1[, 2])))
```
After re-running the simulations with 500 reps and setting a true beta 1 at 0, the minimum p value for Beta 1 above is 0.004, however maximum p value for Beta 1 is 0.99. Due to this range of values, we cannot decisively reject the null.