---
title: "Chapter 6 Question 3"
author: "Allen Church"
date: "November 6, 2019"
output:
  word_document: default
  html_document: default
---

```{r}
library(readxl)
dta <- read_excel("Battleground-65-Final-Dataset.xlsx")
#View(dta)
```

Develop model that addresses relationship between views of civility and views of Trump. Do views about civility affect how democrats vs republicans view president differently?

CIVIL2 = civility
DTID = approval of Trump 
PARTYID = need to convert to dummy variable equals 1 for republicans and 0 for non-republicans
control other variables - no more than 2

Use cross table to view distribution of PARTYID and approximate PARTYID values.
```{r}
library(gmodels)
CrossTable(dta$PARTYID)
```


Use if else statement to create dummy column - if values in PARTYID are greater than 5, assign 1, else assign 0.
```{r}
dta$partyid_dummy <- ifelse(dta$PARTYID>5,1,0)
```

1. Initial difference of means test
```{r}
ols <- lm(dta$DTID ~ dta$partyid_dummy)
summary(ols)
```
The coefficient on partyid_dummy above suggests a mean approval of 2.18851 units for non-Republicans. The model above indicates an increase of 1.58558 in approval units of President Trump associated with Republicans. The intercept combined with the partyid_dummy coefficient suggests the mean approval of President Trump is approximately 3.77409 units for Republicans.

```{r}
ols2 <- lm(dta$DTID ~ dta$partyid_dummy + dta$CIVIL2)
summary(ols2)
```

