---
title: "Chapter 6 Exercise 2"
author: "Allen Church"
date: "11/6/2019"
output: html_document
---

```{r}
require(knitr)
require(haven)
require(AER)

opts_chunk$set(echo = TRUE)
options(digits = 3)

#Add working directory, assign data to lab variable
load("/Users/allenchurch/Ch6_Exercise2_FederalReserve.RData")
```

2a.
Democrat scatterplot
```{r}
#Subset dataset and specify democrat column to equal 1
dem_plot <- subset(dta, dta$democrat==1)

plot(dem_plot$election, dem_plot$FEDFUNDS, main="Quarters and Federal Funds Rate: Democrats",
     xlab="Quarters Since Previous Election", ylab = "FFR (%)", pch = 19, col='blue')
```

Republican Scatterplot
```{r}
#Subset dataset and specify democrat column to equal 0
rep_plot <- subset(dta, dta$democrat==0)

plot(rep_plot$election, rep_plot$FEDFUNDS, main="Quarters and Federal Funds Rate: Republicans",
     xlab="Quarters Since Previous Election", ylab = "FFR (%)", pch = 19, col = "red")
```

2b. Create interaction variable between election and democrat to test whether closeness to elections has the same effect on democrats and republicans.
```{r}
#Create interaction variable by multiplying democrat dummy variable with election
dta$duminteract <- dta$democrat * dta$election
```

Run model with FFR as dependent variable, allowing effect of election variable to vary by party of president
```{r}
olsinteraction <- lm(dta$FEDFUNDS ~ dta$democrat + dta$election + dta$duminteract)
summary(olsinteraction)
```
```{r}
broom::tidy(olsinteraction)
```

2b. i. What change in FFR is associated with a 1 unit increase in election variable when president is republican?
There is a -0.265 change in FFR associated with a 1 unit increase in the election variable when the president is Republican.

2b. ii. What change in FFR is associated with a 1 unit increase in election variable when president is democrat?
There is a 0.558 change in FFR associated with a 1 unit increase in the election variable when the president is Democrat.


2c. Is the effect of election statistically significant under republicans?
```{r}
olsrepublican <- lm(rep_plot$FEDFUNDS ~ rep_plot$election)
summary(olsrepublican)
```

Is the effect of election statistically significant under democrats?
```{r}
olsdemocrat <- lm(dem_plot$FEDFUNDS ~ dem_plot$election)
summary(olsdemocrat)
```

2d. Fitted lines for relationship between election and interest rates - Republicans.
```{r}
plot(rep_plot$election, rep_plot$FEDFUNDS, main="Quarters and Federal Funds Rate: Republicans",
     xlab="Quarters Since Previous Election", ylab = "FFR (%)", pch = 19, col='red')
abline(lm(rep_plot$FEDFUNDS ~ rep_plot$election), col = 'blue')
```

Fitted lines for relationship between election and interest rates - Democrats.
```{r}
plot(dem_plot$election, dem_plot$FEDFUNDS, main="Quarters and Federal Funds Rate: Democrats",
     xlab="Quarters Since Previous Election", ylab = "FFR (%)", pch = 19, col='blue')
abline(lm(dem_plot$FEDFUNDS ~ dem_plot$election), col = 'red' )
```

2e. 
```{r}
olsinteraction_lag <- lm(dta$FEDFUNDS ~ dta$democrat + dta$election + dta$duminteract + dta$lag_FEDFUNDS + dta$inflation)
summary(olsinteraction_lag)
```

