---
title: "Chapter 2 Lab"
author: "Allen Church"
output:
  pdf_document: default
  word_document: default
  html_document: default
---
Load necessary packages
```{r}
library(haven)
```

Set working directory and load lab data
```{r}
lab <- read_dta("Ch2_lab_survey_data.dta")
```


1)	Use the following to create dummy variables for Arlington and Prince William Counties.  How many observations are from each county?

```{r}

#Create dummy variable for Arlington county, selecting corresponding precinct codes with OR operators
lab$Arlington <- (
  lab$precinct == "AR49" | lab$precinct == "AR22" | lab$precinct == "AR2" |
  lab$precinct == "AR18" | lab$precinct == "41" | lab$precinct == "16" |
  lab$precinct == "4" | lab$precinct == "17" | (lab$precinct == "2" &
    
lab$state == 4& !is.na(lab$state))|
  lab$precinct == "31" | lab$precinct == "48")

#Count observations from each county, in this case TRUE corresponds to Arlington
table(lab$Arlington)
```

```{r}
#Create dummy variable for Prince William county, selecting corresponding precinct codes with OR operators
lab$PrinceWilliam1 <- (
  lab$precinct == "PW 101" | lab$precinct == "PW 104" |  lab$precinct == "PW 401" | lab$precinct == "PW101" |
  lab$precinct == "PW104" | lab$precinct == "PW402" | lab$precinct == "PW406"  | lab$precinct == "401" | lab$precinct == "402" | 
(lab$precinct == "104" & lab$state == 4)	)

#Count observations from each county, in this case TRUE corresponds to Prince William
table(lab$PrinceWilliam1)
```

2)	Create dummy variables for each state/DC.  How many observations are in DC, Maryland, Ohio and Virginia?

```{r}
#Create dummy variables for each state
lab$DC <- (lab$state == 1)
lab$Maryland <- (lab$state == 2)
lab$Ohio <- (lab$state == 3)
lab$Virginia <- (lab$state == 4)

#Tabulate observations for each state
table(lab$state)
```

3)	Convert the year_born variable into age.  Be sure to check for and correct for lab errors.  What is the average age of all observations in the lab set?  The minimum and maximum?

```{r}
#Since the survey was taken in 2016, subtract 2016 - year born to obtain age
#Create new age column
lab$age <- 2016 - lab$year_born

#The first summary of the age column shows there is a max age of 152 and 482 NA rows
summary(lab$age)
```

```{r}
#Subset the lab dataframe and exclude values where age is above 100
newdata <- subset(lab, age < 100)

#The summary of the newdata age column shows a new maximum of 95, which is possible
summary(newdata$age)
```

4)	What is the distribution of the gender variable?  Create a male dummy variable and indicate the distribution of this variable. Compare distribution of your male variable to the distribution of the gender variable.

```{r}
#Create male and female variables
lab$male <- (lab$gender == 1)
lab$female <- (lab$gender == 2)
```

Distribution of male variable
```{r}
table(lab$male)
```

Distribution of gender variable. The below table shows that 5 respondents did not identify with the binary gender definition
```{r}
table(lab$gender)
```


5)	Provide descriptive stats for Trump and Clinton feeling thermometer.  Is there anything you need to adjust?

```{r}
#Summarize Clinton feeling thermometer, see there is a max of 200
summary(lab$therm_clinton)
```

```{r}
#Turn values over 100 into NA and summarize again
lab$therm_clinton[lab$therm_clinton > 100] <- NA
summary(lab$therm_clinton)
```


```{r}
#Summarize Trump feeling thermometer
summary(lab$therm_trump)
```

6)	What is the distribution of the education variable?  Is there any adjustment you would need to make if you will use this as a continuous variable in a regression model?

```{r}
#Below shows 7 values for education question
lab1 <- read_dta("Ch2_lab_survey_data.dta")
table(lab1$education)
```

```{r}
#Below we adjust education to exclude the Other response in answer 4, and re-assign the other responses accordingly
#Additionally, the Other response only had 11 values
lab1$education[lab1$education == 4] <- NA
lab1$education[lab1$education == 5] <- 4
lab1$education[lab1$education == 6] <- 5
lab1$education[lab1$education == 7] <- 6

#Generate a new summary table
table(lab1$education)
```

```{r}
#A quick histogram to show the distribution of education
#Below (as well as the table above) shows a low number of responses from 1 - Some high school, which could indicate a bias in sampling, or simply reflect the education of the sample.
hist(lab1$education)
```

