---
title: "Lab2_560"
output: html_document
author: Allen Church
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Name: Allen Church

Chapter 2 Lab Cover Sheet

First load the lab in post_open_refine_clean.dta
load necessary packages
```{r}
library(haven)
```

set working directory (use the path where you saved the file post_open_refine_clean.dta)
For example: here is the folder where I store my lab for this class

load lab

```{r}
lab <- read_dta("Ch2_lab_survey_lab.dta")
```

The codebook for this lab is Codebook for EXIT POLL  survey.docx

1)	Use the following to create dummy variables for Arlington and Prince William Counties.  How many observations are from each county?
Exercise: convert this Stata code to R code

```{r}
lab$Arlington <- (
  lab$precinct == "AR49" | lab$precinct == "AR22" | lab$precinct == "AR2" |
  lab$precinct == "AR18" | lab$precinct == "41" | lab$precinct == "16" |
  lab$precinct == "4" | lab$precinct == "17" | (lab$precinct == "2" &
    
lab$state == 4& !is.na(lab$state))|
  lab$precinct == "31" | lab$precinct == "48")

#count observations from each county
table(lab$Arlington)
```


gen Arlington 		= (precinct == "AR49" | precinct == "AR22" | precinct == "AR2" | precinct == "AR18" | precinct == "41" | /*
*/  precinct == "16" | precinct == "4"   | precinct == "17" | (precinct == "2" & state == 4) | precinct == "31" | precinct == "48" )

gen PrinceWilliam 	= (precinct == "PW 101" | precinct == "PW 104" |  precinct == "PW 401" | precinct == "PW101" | /* 
 	*/ precinct == "PW104" |precinct == "PW402"  |precinct == "PW406"  | precinct == "401" | precinct == "402"  |  /* 
 */  (precinct == "104" & state == 4)	)


2)	Create dummy variables for each state/DC.  How many observations are in DC, Maryland, Ohio and Virginia?

```{r}
lab$DC <- (lab$state ==1)
```




3)	Convert the year_born variable into age.  Be sure to check for and correct for lab errors.  What is the average age of all observations in the lab set?  The minimum and maximum?

```{r}
lab$age[lab$year_born <= 1920] <- NA
lab$age[lab$year_born > 2016] <- NA
summary(lab$age)
View(lab)
```




4)	What is the distribution of the gender variable?  Create a male dummy variable and indicate the distribution of this variable. Compare distribution of your male variable to the distribution of the gender variable.

```{r}
lab$male <- (lab$gender == 1)
lab$female <- (lab$gender == 2)
table(lab$male)
table(lab$gender)

```




5)	Provide descriptive stats for Trump and Clinton feeling thermometer.  Is there anything you need to adjust?

```{r}
summary(lab$therm_clinton)
summary(lab$therm_trump)

lab$therm_clinton[lab$therm_clinton > 100] <- NA
summary(lab$therm_clinton)


```



6)	What is the distribution of the education variable?  Is there any adjustment you would need to make if you will use this as a continuous variable in a regression model?


```{r}
table(lab$education)

lab$education[lab$education == 4] <- NA
lab$education[lab$education == 5] <- 4
lab$education[lab$education == 6] <- 5
lab$education[lab$education == 7] <- 6

table(lab$education)
```





## Including Plots
```{r}
library(haven)
```

```{r}
summary(lab)
```

